# Missing values

---
title: "missing-value"
output: html_notebook
---

```{r}
library(tidyverse)
library(patchwork)
library(ggplot2)

missing_plot <- function(count_or_perc, dataset){
  mycars<-dataset
  colnames(mycars) <- abbreviate(names(mycars), minlength=2)
  missing_patterns <- data.frame(is.na(mycars)) %>%
  group_by_all() %>%
  count(name = "count", sort = TRUE) %>%
  ungroup()

  if (nrow(missing_patterns)>1){
  new_patterns <- data.frame(sapply(missing_patterns, as.character))
  new_patterns[new_patterns == 'TRUE'] <- 1
  new_patterns[new_patterns == 'FALSE'] <- 0
  new_patterns <- data.frame(sapply(new_patterns, as.numeric))
}else{
  new_patterns <- data.frame(t(sapply(missing_patterns, as.character)))
  new_patterns[new_patterns == 'TRUE'] <- 1
  new_patterns[new_patterns == 'FALSE'] <- 0
  new_patterns <- data.frame(t(sapply(new_patterns, as.numeric)))
  }
  
  #top plot
  num_rows_missing <- c()
  for (i in 1:(ncol(new_patterns)-1)) {
    num_rows_missing[i] <- sum(new_patterns[,i]*new_patterns[,'count'])
  }
  top_data <- data.frame(
    columns = colnames(new_patterns)[1:(ncol(new_patterns)-1)],  
    num_rows_missing = num_rows_missing[1:(ncol(new_patterns)-1)]
    )
  top_data <- top_data[order(-1*num_rows_missing), ]
  column_order <- top_data$columns #to help main plot
  top_data$columns <- factor(top_data$columns, levels = top_data$columns)
  top_plot <- ggplot(top_data, aes(x=columns, y=num_rows_missing)) + 
    geom_bar(stat = "identity", fill = 'lightblue') +
    labs(x="varaible", y="num rows missing") + theme(legend.position = "none", axis.text.x = element_text(angle = 45, hjust = 1)) +
    ggtitle("Missing Value Patterns")
  
  #main data, to help right plot
  top_data2 <- top_data[order(-1*num_rows_missing), ]
  main_data <- new_patterns[, 1:(ncol(new_patterns)-1)]
  complete_case <- NULL
  for (i in 1:nrow(main_data)) {
    if (all(main_data[i,]==0)){
      main_data[i,]<-'complete'
      complete_case <- i
    }
  }
  
  #right plot
  if (count_or_perc == 'count'){
    right_data <- data.frame(
    patterns = c(1:nrow(new_patterns)),  
    row_count = new_patterns[, 'count']
    )
    right_data <- right_data %>%
      add_column(ifcomplete='missing')
    for (i in 1:nrow(main_data)) {
      if (all(main_data[i,]=='complete')){
        right_data[i,'ifcomplete']<-'complete'
      }
    }
    right_plot <- ggplot(right_data, aes(x=reorder(patterns, -1*patterns), y=row_count, fill=ifcomplete)) + 
      geom_bar(stat = "identity") +
      scale_fill_manual(values=c('blue', 'lightblue')) +
      labs(x="pattern", y="row_count") + theme(legend.position = "none", axis.text.x = element_text(angle = 45, hjust = 1)) +
      coord_flip()
  } else if (count_or_perc == 'percent') {
     right_data <- data.frame(
      patterns = c(1:nrow(new_patterns)),  
      row_count = new_patterns[, 'count']/sum(new_patterns[, 'count'])
    )
    right_data <- right_data %>%
      add_column(ifcomplete='missing')
    for (i in 1:nrow(main_data)) {
      if (all(main_data[i,]=='complete')){
        right_data[i,'ifcomplete']<-'complete'
      }
    }
    right_plot <- ggplot(right_data, aes(x=reorder(patterns, -1*patterns), y=row_count, fill=ifcomplete)) + 
      geom_bar(stat = "identity") +
      scale_fill_manual(values=c('blue', 'lightblue')) +
      labs(x="pattern", y="row_count_percent") + theme(legend.position = "none", axis.text.x = element_text(angle = 45, hjust = 1)) +
      coord_flip()
  } else {print("input count or percent")}
 
  #main plot
  main_data <- data.frame(cbind(c(1:nrow(new_patterns)), main_data))
  colnames(main_data)[1] <- "pattern"
  main_data <- main_data %>%
    pivot_longer(!pattern, names_to = "variable", values_to = "ifmissing")
  main_data <- data.frame(sapply(main_data, as.character))
  main_data$variable <- fct_relevel(main_data$variable, column_order)
  main_data$pattern <- fct_relevel(main_data$pattern, as.character(c(1:nrow(new_patterns))))
  
  main_plot <- ggplot(main_data, aes(x=variable, y=pattern, fill=ifmissing))+
    scale_y_discrete(limits = rev) +
    scale_fill_manual(values = c('1'='purple', '0'='grey', 'complete'='darkgrey'))+
    geom_tile(color = "white") +
    theme(legend.position = "none", axis.text.x = element_text(angle = 45, hjust = 1)) +
    labs(x='variable',y='pattern')
  
  #patchwork
  x <- as.integer(ncol(mycars)/2)
  y <- nrow(missing_patterns)-as.integer(complete_case)+1
  main_plot <- main_plot + geom_text(aes(x,y,label='complete case'), size=4)

  top_plot + plot_spacer() + main_plot + right_plot + 
    plot_layout(ncol = 2, nrow = 2, widths = c(4, 1), heights = c(1, 4))
}
```


```{r}
nursing = read.csv("./nursinghouse_data.csv")
missing_plot('count', nursing)
```

We here investigate missing values in 'new_nursing.csv', the processed ‘COVID-19 Nursing Home Data’. By aggragting data with nearby dates and select dates with sufficient data, we obatin a dataset with 91892 records for around 15 thousand nursing homes with 6 main times points. According to the missing value plots shown above, We can clearly see that there are not many rows which have lots of missing values. The number of the row pattern with the most missing values is less than 1500 or 1%, which is quite small compared with the complete case row. The numbers of all other row patterns are also very small. This preprocessing sucessfully extract an informative subset from the original dataset and supports our analysis siginifcantly.

During the analysis, we further address the remaining missing values in this subset. The top four columns with the most missing values(T.R.C, T.R.CO, N, T.N) are Total Residents Confirmed Cases, Total Residents Covid-19 Death, Number of All Beds, Total Number of Occupied Beds. They are all numerical values. All other columns which have missing values are also numerical values. Since we have state information for nursing houses, we replace all missing values of a column by the average of other values which are in the same state.